---
title: "hw2_Xuan_Huang"
author: "Xuan Huang"
date: "2022-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r needed_library}
library(dplyr)
library(data.table)
library(tidyverse)
library(data.table)
library(gtools)
library(ggpmisc)
library(graphics)
library(leaflet)
```

# Data Wrangling


```{r}
# download data
if (!file.exists("chs_individual.csv")) {
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv", 
    destfile = "chs_individual.csv", 
    method="libcurl", 
    timeout = 60
    )
}
if (!file.exists("chs_regional.csv")) {
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv", 
    destfile = "chs_regional.csv", 
    method="libcurl", 
    timeout = 60
    )
}


```

```{r}
# read data
chsInd <- fread("chs_individual.csv")
chsReg <- fread("chs_regional.csv")
head(chsInd)
head(chsReg)
```

### 1. After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches.

```{r}
dim(chsInd)
dim(chsReg)

chsMerge <- merge(
  x = chsInd,
  y = chsReg,
  by.x = "townname",
  by.y = "townname",
  all.x = T,
  all.y = F
)
dim(chsMerge)

# After merging the data, the rows of merged dataset is equal to the rows of 
# individual CHS dataset

# To make sure there is no duplicates
chsMerge <- unique(chsMerge)

# Impute data using the average within the variables “male” and “hispanic” for
# missing value
#summary(chsMerge)
chsMale_His <- chsMerge %>% 
  filter(male == 1 & hispanic == 1)

chsMerge[is.na(agepft), agepft := round(mean(chsMale_His$agepft, na.rm = T), digits = 3)]
chsMerge[is.na(height), height := round(mean(chsMale_His$height, na.rm = T), digits = 0)]
chsMerge[is.na(weight), weight := round(mean(chsMale_His$weight, na.rm = T), digits = 0)]
chsMerge[is.na(bmi), bmi := round(mean(chsMale_His$bmi, na.rm = T), digits = 3)]
chsMerge[is.na(asthma), asthma := round(mean(chsMale_His$asthma, na.rm = T), 1)]
chsMerge[is.na(father_asthma), father_asthma := round(mean(chsMale_His$father_asthma, na.rm = T), 1)]
chsMerge[is.na(mother_asthma), mother_asthma := round(mean(chsMale_His$mother_asthma, na.rm = T), 1)]
chsMerge[is.na(wheeze), wheeze := round(mean(chsMale_His$wheeze, na.rm = T), 1)]
chsMerge[is.na(hayfever), hayfever := round(mean(chsMale_His$hayfever, na.rm = T), 1)]
chsMerge[is.na(allergy), allergy := round(mean(chsMale_His$allergy, na.rm = T), 1)]
chsMerge[is.na(educ_parent), educ_parent := round(mean(chsMale_His$educ_parent, na.rm = T), 1)]
chsMerge[is.na(smoke), smoke := round(mean(chsMale_His$smoke, na.rm = T), 1)]
chsMerge[is.na(gasstove), gasstove := round(mean(chsMale_His$gasstove, na.rm = T), 1)]
chsMerge[is.na(fev), fev := round(mean(chsMale_His$fev, na.rm = T), digits = 1)]
chsMerge[is.na(fvc), fvc := round(mean(chsMale_His$fvc, na.rm = T), 1)]
chsMerge[is.na(mmef), mmef := round(mean(chsMale_His$mmef, na.rm = T), digits = 1)]

#summary(chsMerge)
```

### 2. Create a new categorical variable named “obesity_level” using the BMI measurement 

```{r}
# create categorical variable
chsMerge$obesity_level <- cut(chsMerge$bmi,
                              breaks = c(0, 14, 22, 24, Inf),
                              labels = c("underweight", "normal", "overweight", "obese")
                              )

# create a summary table for bmi level
chsMerge %>%
  group_by(obesity_level) %>%
  summarise(
    min = min(bmi),
    max = max(bmi),
    count = n()
  )

```

### 3. Create another categorical variable named “smoke_gas_exposure” 

```{r}
#“smoke_gas_exposure”  summarizes “Second Hand Smoke” and “Gas Stove.”
summary(chsMerge$smoke)
summary(chsMerge$gasstove)

# there are 4 possible combination of second hand smoke status and gas stove status
chsMerge <- 
  chsMerge %>%
  mutate(
    smoke_gas_exposure = case_when(
      smoke == 0 & gasstove == 0 ~ "neiher",
      smoke == 1 & gasstove == 0 ~ "only smoke",
      smoke == 0 & gasstove == 1 ~ "only gas stove",
      smoke == 1 & gasstove == 1 ~ "both"
    )
  )

table(chsMerge$smoke_gas_exposure)
```

### 4. Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”

```{r}
# summary table by town
chsMerge %>% 
  group_by(townname) %>%
  summarise(
    count = n(),
    average_fev = mean(fev),
    sd_fev = sd(fev),
    prop_asthma = mean(asthma),
    sd_asthma = sd(asthma)
            )
```
```{r}
# summary table by sex
chsMerge %>% 
  group_by(male) %>%
  summarise(
    count = n(),
    average_fev = mean(fev),
    sd_fev = sd(fev),
    prop_asthma = mean(asthma),
    sd_asthma = sd(asthma)
            )
```

```{r}
# by obesity level
chsMerge %>% 
  group_by(obesity_level) %>%
  summarise(
    count = n(),
    average_fev = mean(fev),
    sd_fev = sd(fev),
    prop_asthma = mean(asthma),
    sd_asthma = sd(asthma)
            )
```

```{r}
# by smoke_gas_exposure
chsMerge %>% 
  group_by(smoke_gas_exposure) %>%
  summarise(
    count = n(),
    average_fev = mean(fev),
    sd_fev = sd(fev),
    prop_asthma = mean(asthma),
    sd_asthma = sd(asthma)
            )
```


# Looking at the Data (EDA)

### The primary questions of interest are: 
- 1. What is the association between BMI and FEV (forced expiratory volume)? 
**People with a higher BMI level will have a higher forced expiratory volume.**
- 2. What is the association between smoke and gas exposure and FEV? 
**There is no clear association between smoke and gas exposure and FEV. People contact with only smoke exposure will have higher FEV.**
- 3. What is the association between PM2.5 exposure and FEV?
**The average forced expiratory volume for people living in a given town will decrease with the increase of PM2.5 concentration.**

### EDA checklist

```{r}
# check the dimensions and headers and footers
dim(chsMerge)
head(chsMerge)
tail(chsMerge)
```
```{r}
#check the variable types
str(chsMerge)
```

```{r}
# Take a closer look at some/all of the variables
# here, our key variables are bmi, fev, smoke_gas_exposure, PM2.5
summary(chsMerge$bmi)
summary(chsMerge$fev)
table(chsMerge$smoke_gas_exposure)
summary(chsMerge$pm25_mass)
```

**After taking a closer look of key variables, there is no unreasonable value of key variables. So we don't need to clean our data.**

### 1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.

```{r}
chsMerge %>%
  ggplot(mapping = aes(x = bmi, y = fev, color = townname)) +
  geom_point() +
  geom_smooth(method = "lm", color = "Black") +
  labs(title = "BMI vs FEV by Twon",x = "BMI", y = "FEV") +
  facet_wrap(~ townname)
```

**As shown of these scatter plots, there is a positive relationship between BMI and FEV. With the increasing of BMI, forced expiratory volume also increased. For the different towns, this correlation can be strong or weak. So we can speculate that the correlation between BMI and FEV can vary by towns.**


### 2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.

```{r}
# Stacked histograms of FEV by BMI category
chsMerge %>%
  ggplot(aes(x = fev, fill = obesity_level)) +
  geom_histogram() +
  scale_fill_manual(values = c("red", "blue", "darkgreen", "orange")) +
  labs(title = "FEV by Obesity level",x = "FEV", y = "Count", fill = "Obesity level")
```

```{r}
# FEV by smoke/gas exposure
chsMerge %>%
  ggplot(aes(x = fev, fill = smoke_gas_exposure)) +
  geom_histogram() +
  scale_fill_manual(values = c("red", "blue", "darkgreen", "orange")) +
  labs(title = "FEV by smoke/gas exposure", x = "FEV", y = "Count", fill = "smoke/gas exposure")
```

**From the stacked histogram for obesity level, the curve of overweight and obese is right-skewed, which means the overweight or obese people are more likely to have higher FEV. The sample size of overweight and obese is much lower than underweight and normal categories, so the curve of overweight and obese seem flatter. And there is significantly high bin at about 2200.**

**From the stacked histogram for smoke/gas exposure, there is no significant skewed in four categories. Four categories have similar means. And also when fev is around 2200, the count of four categories get maximum. Except only smoke exposure category, other three categories have abundant data. Low sample size can affect our conclusion about only smoke exposure category.**

### 3. Barchart of BMI by smoke/gas exposure.

```{r}
chsMerge %>%
  ggplot(aes(x = obesity_level, fill = smoke_gas_exposure)) +
  geom_bar() + 
  scale_fill_manual(values = c("red", "blue", "darkgreen", "orange")) +
  labs(title = "Barchart of BMI by smoke/gas exposure", x = "Obesity level", y = "Count", fill = "Smoke/Gas exposure")
```

**From the barchart, the "both" exposure takes up the most proportion for each obesity level. And the normal BMI level have the most abundant sample size. So analysis of the effect of smoke/gas exposure in normal people will get the most powerful result. The "only smoke" exposure exists a small amount in four obesity levels. So it will be hard to analyze the effect of "only smoke" exposure.**

### 4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.

```{r}
chsMerge %>%
  ggplot(mapping = aes(y = fev, x = obesity_level),
         fun = median,
         fun.min = min,
         fun.max = max
         ) +
  stat_summary() +
  labs(title = "Statistical Summary of FEV by BMI", y = "FEV", x = "Obesity_level")

```

```{r}
chsMerge %>%
  ggplot(mapping = aes(y = fev, x = smoke_gas_exposure),
         fun = median,
         fun.min = min,
         fun.max = max
         ) +
  stat_summary() +
  labs(title = "Statistical Summary of FEV by Smoke/Gas exposure", y = "FEV", x = "Smoke/Gas exposure")
```

**The statistical summary graph of FEV by BMI shown that with the increase of BMI level, people's average FEV value increase. Obese people have the most forced expiratory volume.**

**From the statistical summary graph of FEV by gas/smoke exposure, people in only smoke exposure have the most forced expiratory volume. And the range of FEV in only smoke exposure is also the largest one.**


### 5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

```{r}
temp.pal <- colorFactor(palette ='viridis', domain = chsMerge$pm25_mass)

leaflet(chsMerge[!is.na(pm25_mass)]) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
    lat = ~lat, lng=~lon,
    label = chsMerge$townname, color = ~ temp.pal(pm25_mass),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=temp.pal, values = chsMerge$pm25_mass,
          title='Concentrations of PM2.5 mass in each of the CHS communities', opacity=1)
```

**From the leaflet plot, we can know the geographical information of concentration of PM2.5 mass.The highest PM2.5 concentrations happened in Mira Loma. The lowest PM2.5 concentrations happened in the Atascadero. The air quality of Pasadena is not very good. The coast area of Santa Barbara have a good air quality.**


### 6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.

```{r}
# Both PM2.5 and FEV are numeric variable, so we can create a scatter plot with a linear regression model
chsMerge %>%
  ggplot(mapping = aes(x = pm25_mass, y = fev)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  labs(title = "Relationship between PM2.5 and FEV")
```

```{r}
# Based on the previous scatter plot, we cannot find a obviously association.
# Because PM2.5 mass is a town_level variable, we should calculate the typical 
# FEV for each town and research the relationship between the town level FEV and
# twon level PM2.5 mass

chsMerge_town <- chsMerge %>%
  group_by(townname) %>%
  mutate(fev_town = mean(fev))

chsMerge_town %>%
  ggplot(mapping = aes(x = pm25_mass, y = fev_town, color = factor(townname))) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  labs(title = "Relationship between PM2.5 and FEV", x = "PM2.5 Concentration for each Town", y = "Average FEV for each Town")
  
```

**Based on the linear regression model, the average forced expiratory volume for people living in a given town will decrease with the increase of PM2.5 concentration.**

### The primary questions of interest are: 
- 1. What is the association between BMI and FEV (forced expiratory volume)? 
**People with a higher BMI level will have a higher forced expiratory volume.**
- 2. What is the association between smoke and gas exposure and FEV? 
**There is no clear association between smoke and gas exposure and FEV. People contact with only smoke exposure will have higher FEV.**
- 3. What is the association between PM2.5 exposure and FEV?
**The average forced expiratory volume for people living in a given town will decrease with the increase of PM2.5 concentration.**


