---
title: "HW3"
author: "Xuan Huang"
date: "2022-11-05"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(xml2)
library(stringr)
library(tidyverse)
library(tibble)
library(dplyr)
library(ggplot2)
library(tidytext)
```


# APIs
- Using the NCBI API, look for papers that show up under the term “sars-cov-2 trial vaccine.” Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r}
# Download web, and count the numbers
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]/span")

counts <- as.character(counts)
```


```{r}
stringr::str_extract(counts, "[0-9,]+")

# we found 4009 papers
```

- Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r}
# download details for first 250, get those IDs
query_ids <- GET(
      url   = "https://eutils.ncbi.nlm.nih.gov/",
      path = "entrez/eutils/esearch.fcgi",
      query = list(db       = "pubmed",
                   term     = "sars-cov-2 trial vaccine",
                   rettype = "abstract",
                   retmax   = 250)
    )
ids <- httr::content(query_ids)
ids <- as.character(ids)
ids <- stringr::str_extract_all(ids, "<Id>[0-9]+</Id>")[[1]]
ids <- stringr::str_remove_all(ids, "<Id>|</Id>")
```


- As we did in lab 7. Create a dataset containing the following:

Pubmed ID number,
Title of the paper,
Name of the journal where it was published,
Publication date, and
Abstract of the paper (if any).

```{r}
# create a dataset contain those information
publications <- GET(
  url = "https://eutils.ncbi.nlm.nih.gov/",
  path = "entrez/eutils/efetch.fcgi",
  query = list(
    db = "pubmed",
    id = paste(ids, collapse = ","),
    retmax = 250,
    rettype = "abstract"
    ))
publications <- httr::content(publications)
publications_txt <- as.character(publications)

pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
    
# From Lab 7, and the other variables are modeled after the next three lines of code
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>")
abstracts <- str_replace_all(abstracts, "[[:space:]]+", " ")

# For article titles...
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")
  
# For journal titles...
journals <- str_extract(pub_char_list, "<Title>[[:print:][:space:]]+</Title>")
journals <- str_remove_all(journals, "</?[[:alnum:]- =\"]+>") 
journals <- str_replace_all(journals, "[[:space:]]+", " ")
  
# For publciation dates
dates <- str_extract(pub_char_list, "<PubDate>[[:print:][:space:]]+</PubDate>")
dates <- str_remove_all(dates, "</?[[:alnum:]- =\"]+>")
dates <- str_replace_all(dates, "[[:space:]]+", " ")

database <- data.frame(
      PubMedID = ids,
      Title = titles,
      Journal = journals,
      PubDate = dates,
      Abstract = abstracts
    )

## show the table
knitr::kable(database[1:20, ], caption = "Twenty papers about SARS-CoV-2 Trial Vaccines")
```


# Text Mining

A new dataset has been added to the data science data repository https://github.com/USCbiostats/data-science-data/tree/master/03_pubmed. The dataset contains 3241 abstracts from articles across 5 search terms. Your job is to analyse these abstracts to find interesting insights.

1. Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?

```{r}
# Read in data
if (!file.exists("pubmed.csv")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv", "pubmed.csv", method="libcurl", timeout = 60)
}
pubmed <- read.csv("pubmed.csv")
pubmed <- as_tibble(pubmed)

# With stop words
pubmed %>%
  unnest_tokens(output = word, input = abstract) %>%
  count(word, sort = TRUE) %>%
  top_n(20) %>%
  ggplot(aes(x = n, y = fct_reorder(word,n))) +
  geom_col() +
  labs(x = "Numbers", y = "Word", title = "Top 20 Tokens (with stop words)")




```

From the summary of occurrence number， most of them are stop words, which are not helpful for us to infer the subject of the paper.


```{r}
# Without stop words
pubmed %>%
  unnest_tokens(output = word, input = abstract) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = "word") %>%
  filter(!grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(20) %>%
  ggplot(aes(x = n, y = fct_reorder(word,n))) +
  geom_col() +
  labs(x = "Numbers", y = "Word", title = "Top 20 Tokens (without stop words)")

pubmed %>%
  unnest_tokens(output = word, input = abstract) %>%
  anti_join(stop_words, by = "word") %>%
  group_by(term) %>%
  count(word, sort = TRUE) %>%
  top_n(5) %>%
  knitr::kable(caption = "Top 5 Tokens for Each Term")
```
After cleaning the stop words, the top 5 common tokens were "covid", "patients", "cancer", and "prostate", "disease".


2. Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.

```{r}
pubmed %>%
  unnest_ngrams(output = bigram, input = abstract, n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>% 
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>% 
  unite(bigram, word1, word2, sep = " ") %>% 
  count(bigram, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = n, y = fct_reorder(bigram,n))) +
  geom_col() +
  labs(x = "Numbers", y = "Bigram", title = "Top 10 Bigrams")
```

We find that the occurrence of "covid 19" is much higher than other bigrams.


3. Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”) What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?

```{r}
#tokens with stop words and number
pubmed %>%
  unnest_tokens(token, abstract) %>%
  count(token, term) %>%
  bind_tf_idf(token, term, n) %>%
  arrange(desc(tf_idf)) %>%
  top_n(5, tf_idf)

#token without stop words and number
pubmed %>%
  unnest_tokens(token, abstract) %>%
  anti_join(stop_words, by = c("token" = "word")) %>%
  filter(!grepl(pattern = "^[0-9]+$", x = token)) %>%
  count(token, term) %>%
  bind_tf_idf(token, term, n) %>%
  arrange(desc(tf_idf)) %>%
  top_n(5, tf_idf)

#bi-grams with stop words and number
pubmed %>%
  unnest_ngrams(bigram, abstract, n=2) %>%
  count(bigram, term) %>%
  bind_tf_idf(bigram, term, n) %>%
  arrange(desc(tf_idf)) %>%
  top_n(5, tf_idf)
```
```

The 5 common tokens without stop words in question 1 are "covid", "patients", "cancer", and "prostate", "disease".
Here, 5 common tokens with highest tf-idf value are "covid", "prostate", "eclampsia","preeclampsia", "meningitis".
We find there is a difference in "eclampsia","preeclampsia", "meningitis". Those three common tokens with highest tf-idf value are specific diseases.

For bigrams, the answer of question 1 and question 3 are same.
