---
title: "lab5_Xuan_Huang"
author: "Xuan Huang"
date: "2022-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r need library}
library(lubridate)
library(data.table)
library(dtplyr)
```

```{r}
# Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

```{r}
# merge met data
dat <- fread("met_all.gz")
dat <- dat[temp > -17]
str(stations)
str(dat)

met <- merge(
# Data
 x = dat, 
 y = stations, 
# List of variables to match
 by.x = "USAFID",
 by.y = "USAF", 
# Which obs to keep?
 all.x = TRUE, 
 all.y = FALSE
 ) 
str(met)

```

### Question 1: Representative station for the US
What is the median station in terms of temperature, wind speed, and atmospheric pressure? Look for the three weather stations that best represent continental US using the quantile() function. Do these three coincide?

*Computer mean temperature, wind speed, and atmospheric pressure for each weather station, and pick the weather station with the average value closest to the median*
```{r}
# calculate mean for each station
met_mean <- met[, .(
  temp = mean(temp, na.rm = T),
  wind.sp = mean(wind.sp, na.rm = TRUE),
  atm.press = mean(atm.press, na.rm = TRUE)
), by = USAFID]

str(met_mean)
summary(met_mean$atm.press)
dim(met_mean)
length(unique(met$USAFID))
```

```{r}
# calculate median

met_median <- met_mean[,.(
  temp_50 = quantile(temp, probs = 0.5, na.rm = TRUE),
  wind.sp_50 = quantile(wind.sp, probs = 0.5, na.rm = TRUE),
  atm.press_50 = quantile(atm.press, probs = 0.5, na.rm = TRUE)
)]

str(met_median)
```

```{r}
met_mean[, temp_dist := abs(temp - met_median$temp_50)]

median_temp_station <- met_mean[order(temp_dist)][1:3]
median_temp_station

met_mean[which.min(temp_dist)]
```

### Question 2: Representative station per state
Just like the previous question, you are asked to identify what is the most representative, the median, station per state. This time, instead of looking at one variable at a time, look at the euclidean distance. If multiple stations show in the median, select the one located at the lowest latitude.

```{r}
table(met$STATE)

met_mean_state <- met[, .(
  temp = mean(temp, na.rm = T),
  wind.sp = mean(wind.sp, na.rm = TRUE),
  atm.press = mean(atm.press, na.rm = TRUE)
), by = .(USAFID, STATE)]

str(met_mean_state)

met_median_state <- met_mean_state[,.(
  temp_50 = quantile(temp, probs = 0.5, na.rm = TRUE),
  wind.sp_50 = quantile(wind.sp, probs = 0.5, na.rm = TRUE),
  atm.press_50 = quantile(atm.press, probs = 0.5, na.rm = TRUE)
), by = STATE]

str(met_median_state)
str(met_mean_state)

met_state <- merge(
# Data
 x = met_mean_state, 
 y = met_median_state, 
# List of variables to match
 by.x = "STATE",
 by.y = "STATE", 
# Which obs to keep?
 all.x = TRUE, 
 all.y = FALSE
 ) 
str(met_state)

```

```{r}
class(met_state)
#met_state <- met_state[, c(temp_50_dist := abs(temp - temp_50), wind_50_dist := abs(wind.sp - wind.sp_50))] # := just can have one?
met_state <- met_state[, temp_50_dist := abs(temp - temp_50)]
met_state <- met_state[, wind_50_dist := abs(wind.sp - wind.sp_50)]
met_state <- met_state[, eucdist := sqrt(temp_50_dist^2 + wind_50_dist^2)]
str(met_state)

met_state[order(eucdist)][1:5]


```



